plyranges: a grammar of genomic data transformation

### 2018-10-24 paper.Rmd
final readable version

---
title: "plyranges: a grammar of genomic data transformation" 
author:
  - name: Stuart Lee \footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}
    email: stuart.lee1@monash.edu
  - name: Dianne Cook $^{\ref{monash}}$
    email: di.cook@monash.edu 
  - name: Michael Lawrence \footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author \href{mailto:michael.lawerence@gene.com}{michael.lawerence@gene.com} } 
    email: lawrence.michael@gene.com
abstract: "There is a cognitive load placed on users in learning a data abstraction from the Bioconductor project and understanding its appropriate use. Users must navigate these abstractions to perform a genomic analysis task, when a single data abstraction, a GRanges object will suffice. By recognizing that the GRanges class follows 'tidy' data principles, we create a grammar of genomic data transformation, defining verbs for performing actions on and between genomic interval data and providing a way of performing common data analysis tasks through a coherent interface to existing Bioconductor infrastructure. We implement this grammar as a Bioconductor/R package called \texttt{plyranges}."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis" 
bibliography: references.bib 
output: 
  pdf_document:
    template: svm-latex-ms.tex 
    keep_tex: true 
    fig_caption: true
    latex_engine: pdflatex 
    citation_package: biblatex 
fontfamily: mathpazo
---

```{r setup, include = FALSE}
library(tibble)
library(knitr)
opts_chunk$set(message = FALSE, 
               warning = FALSE, 
               comment = "#>", 
               fig.width = 5, 
               fig.height = 3, 
               fig.align = "center",
               fig.path = "diagrams")
# for peak figures
library(ggbio)

@@

\begin{table}[!htbp]
\centering
\begin{tabular}{|l|l|p{6cm}|}
  \hline
  & Verb &  Description \\ 
  \hline
   & \textbf{\emph{summarize()}} & aggregate over column(s) \\ 
   Aggregate & \emph{disjoin\_ranges()} & aggregate column(s) over the union of end coordinates \\
   &  \emph{reduce\_ranges()} & aggregate column(s) by merging overlapping and neighboring ranges \\
   \hline
   &  \textbf{\emph{mutate()}} & modifies any column \\
   & \textbf{\emph{select()}} & select columns \\
  Modify (Unary) & \textbf{\emph{arrange()}} & sort by columns \\
   & \emph{stretch()} & extend range by fixed amount \\
   &  \emph{shift\_(direction)} & shift coordinates \\
   & \emph{flank\_(direction)} & generate flanking regions \\
   & \emph{\%intersection\% } & row-wise intersection \\
   & \emph{\%union\%} & row-wise union \\
   & \emph{compute\_coverage} & coverage over all ranges \\
  Modify (Binary) &  \emph{\%setdiff\%} & row-wise set difference \\
   & \emph{between()} & row-wise gap range \\
   & \emph{span()} & row-wise spanning range \\
   \hline
    & \emph{join\_overlap\_*()} & merge by overlapping ranges \\
    & \emph{join\_nearest} & merge by nearest neighbor ranges \\
    & \emph{join\_follow} & merge by following ranges \\
    Merge & \emph{join\_precedes} & merge by preceding ranges \\
    & \emph{union\_ranges} & range-wise union \\
    & \emph{intersect\_ranges} & range-wise intersect \\
    & \emph{setdiff\_ranges} & range-wise set difference \\
    & \emph{complement\_ranges} & range-wise union \\
  \hline
   & \emph{anchor\_direction()} & fix coordinates at direction \\
  Operate & \textbf{\emph{group\_by()}} & partition by column(s)  \\ 
   & \emph{group\_by\_overlaps()} & partition by overlaps \\
   \hline
   & \textbf{\emph{filter()}} & subset rows \\
  Restrict & \emph{filter\_by\_overlaps()} & subset by overlap \\
    & \emph{filter\_by\_non\_overlaps()} & subset by no overlap \\
   \hline
\end{tabular}
\caption{Overview of the \texttt{plyranges} grammar. The core verbs are
briefly described and categorized into one of the following higher level 
categories: aggregate, modify, merge, operate, or restrict. A verb is given bold text if
its origin is from the \texttt{dplyr} grammar.}\label{tab:grammar}
\end{table}

@@

The design of `plyranges` adheres to well understood principles of
language and API design: cognitive consistency, cohesion,
expressiveness and endomorphism \cite{Green1996-qg}. 
To varying degrees, these principles also underlie the
design of `dplyr` and the Bioconductor infrastructure.

@@

The `plyranges` package can be obtained via the Bioconductor project
website [https://bioconductor.org](https://bioconductor.org) or
accessed via Github 
[https://github.com/sa-lee/plyranges](https://github.com/sa-lee/plyranges).

# Methods
## Data Availability

## Software Versions
To produce the workflows as described in results section we used 
R version 3.5 with the development version of `plyranges` (1.1.10)   
and the `BioStrings` package (2.49.0) installed \cite{r-core, plyranges-zenodo, R-biostrings}.

All code required to reproduce this article is available at 
https://github.com/sa-lee/plyranges-paper.

@@

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall 
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages
for the code review they performed and users who have submitted feedback
and pull requests. This article was written with `knitr` \cite{R-knitr} and 
the figures were made with `ggbio` \cite{R-ggbio}.

>>>>

---
title: "plyranges: a grammar of genomic data transformation" 
author:
  - name: Stuart Lee \footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}
    email: stuart.lee1@monash.edu
  - name: Dianne Cook $^{\ref{monash}}$
    email: di.cook@monash.edu 
  - name: Michael Lawrence \footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author \href{mailto:michael.lawerence@gene.com}{michael.lawerence@gene.com} } 
    email: lawrence.michael@gene.com
abstract: "There is a cognitive load placed on users in learning a data abstraction from the Bioconductor project and understanding its appropriate use. Users must navigate these abstractions to perform a genomic analysis task, when a single data abstraction, a GRanges object will suffice. By recognizing that the GRanges class follows 'tidy' data principles, we create a grammar of genomic data transformation, defining verbs for performing actions on and between genomic interval data and providing a way of performing common data analysis tasks through a coherent interface to existing Bioconductor infrastructure. We implement this grammar as a Bioconductor/R package called \\texttt{plyranges}."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis" 
bibliography: references.bib 
output: 
  pdf_document:
    template: svm-latex-ms.tex 
    keep_tex: true 
    fig_caption: true
    latex_engine: pdflatex 
    citation_package: biblatex 
fontfamily: mathpazo
---

```{r setup, include = FALSE}
library(tibble)
library(knitr)
opts_chunk$set(message = FALSE, 
               warning = FALSE, 
               comment = "#>", 
               fig.width = 5, 
               fig.height = 3, 
               fig.align = "center",
               fig.path = "./diagrams/")
# for peak figures
library(ggbio)

@@

\begin{table}[!htbp]
\centering
\begin{tabular}{|l|l|p{6cm}|}
  \hline
  & Verb &  Description \\ 
  \hline
   & \textbf{\emph{summarize()}} & aggregate over column(s) \\ 
   Aggregate & \emph{disjoin\_ranges()} & aggregate column(s) over the union of end coordinates \\
   &  \emph{reduce\_ranges()} & aggregate column(s) by merging overlapping and neighboring ranges \\
   \hline
   &  \textbf{\emph{mutate()}} & modifies any column \\
   & \textbf{\emph{select()}} & select columns \\
  Modify (Unary) & \textbf{\emph{arrange()}} & sort by columns \\
   & \emph{stretch()} & extend range by fixed amount \\
   &  \emph{shift\_(direction)} & shift coordinates \\
   & \emph{flank\_(direction)} & generate flanking regions \\
   & \emph{\%intersection\% } & row-wise intersection \\
   & \emph{\%union\%} & row-wise union \\
   & \emph{compute\_coverage} & coverage over all ranges \\
  Modify (Binary) &  \emph{\%setdiff\%} & row-wise set difference \\
   & \emph{between()} & row-wise gap range \\
   & \emph{span()} & row-wise spanning range \\
   \hline
    & \emph{join\_overlap\_*()} & merge by overlapping ranges \\
    & \emph{join\_nearest} & merge by nearest neighbor ranges \\
    & \emph{join\_follow} & merge by following ranges \\
    Merge & \emph{join\_precedes} & merge by preceding ranges \\
    & \emph{union\_ranges} & range-wise union \\
    & \emph{intersect\_ranges} & range-wise intersect \\
    & \emph{setdiff\_ranges} & range-wise set difference \\
    & \emph{complement\_ranges} & range-wise set complement \\
  \hline
   & \emph{anchor\_direction()} & fix coordinates at direction \\
  Operate & \textbf{\emph{group\_by()}} & partition by column(s)  \\ 
   & \emph{group\_by\_overlaps()} & partition by overlaps \\
   \hline
   & \textbf{\emph{filter()}} & subset rows \\
  Restrict & \emph{filter\_by\_overlaps()} & subset by overlap \\
    & \emph{filter\_by\_non\_overlaps()} & subset by no overlap \\
   \hline
\end{tabular}
\caption{Overview of the \texttt{plyranges} grammar. The core verbs are
briefly described and categorized into one of the following higher level 
categories: aggregate, modify, merge, operate, or restrict. A verb is given bold text if
its origin is from the \texttt{dplyr} grammar.}\label{tab:grammar}
\end{table}

@@

The design of `plyranges` adheres to well understood principles of
language and API design: cognitive consistency, cohesion,
endomorphism and expressiveness \cite{Green1996-qg}. 
To varying degrees, these principles also underlie the
design of `dplyr` and the Bioconductor infrastructure.

@@

@@

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall 
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages
for the code review they performed and users who have submitted feedback
and pull requests.

## Availabliity of data and Materials

The `plyranges` package is open source under an Artistic 2.0 license 
\cite{R-plyranges-z}. 
The  version (1.1.10) used in the paper is archived under DOI: [10.5281/zenodo.1469841](https://doi.org/10.5281/zenodo.1469841). The software 
can be obtained via the Bioconductor project website 
[https://bioconductor.org](https://bioconductor.org) or accessed via Github 
[https://github.com/sa-lee/plyranges](https://github.com/sa-lee/plyranges).

To produce the workflows as described in results section we used 
R version 3.5 with the development version of `plyranges` (1.1.10)   
and the `BioStrings` package (2.49.0) installed \cite{r-core, R-plyranges-z, R-biostrings}.

This article was written with `knitr` \cite{R-knitr} and 
the figures were made with `ggbio` \cite{R-ggbio}. All code required to 
reproduce this article is available at https://github.com/sa-lee/plyranges-paper.

### 2018-10-24 paper.tex
final readable version

% highlight changes
\usepackage{soul}

@@

\title{plyranges: A grammar of genomic data transformation  }

\author{\Large Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Dianne Cook \(^{\ref{monash}}\)\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author}\vspace{0.05in} \newline\normalsize\emph{}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia} () and Dianne Cook \(^{\ref{monash}}\) () and Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author} ()},
             pdfkeywords = {Bioconductor, Grammar, Genomes, Data Analysis},  
            pdftitle={plyranges: A grammar of genomic data transformation},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Dianne Cook \(^{\ref{monash}}\)} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author}} \hskip 15pt \emph{\small }

@@

\noindent The Bioconductor project provides many interoperable data abstractions
for analyzing high-throughput genomics experiments; however implementing
a typical genomic workflow with Bioconductor requires learning these
abstractions and understanding them at an integrative level. This places
a large cognitive burden on the user, especially for non-programmers. To
reduce this burden we have created a grammar of genomic data
transformation that operates on a single, central Bioconductor data
structure, GRanges, which naturally represents genomic intervals and
their associated measurements. The grammar defines verbs for performing
actions on and between genomic interval data through a simplified,
coherent interface to existing Bioconductor infrastructure, resulting in
fluent analysis workflows. We have implemented this grammar as an
R/Bioconductor package called \texttt{plyranges}.

@@

Recently, there has been a concerted, community effort to standardize R
data structures and workflows around the notion of tidy data
\cite{Wickham2014-jc}. A tidy dataset is defined as a tabular data
structure that has observations as rows and columns as variables, and
all measurements pertain to a single observational unit. The tidy data
pattern is useful because it allows us to see how the data relate to the
design of an experiment and the variables measured. The \texttt{dplyr}
package \cite{Wickham2017-dplyr} defines an application programming
interface (API) that maps notions from query languages such as SQL to
verbs that act on tidy data. These verbs can be composed together on one
or more tidy datasets with the pipe operator from the \texttt{magrittr}
package \cite{R-magrittr}. Taken together these features enable a user
to write human readable analysis workflows.

@@

{\centering \includegraphics{paper_files/figure-latex/peak-viz-1}

@@

{\centering \includegraphics{paper_files/figure-latex/titv-viz-1}

@@

The design of \texttt{plyranges} adheres to well understood principles
of language and API design: cognitive consistency, cohesion,
expressiveness and endomorphism \cite{Green1996-qg}. To varying degrees,
these principles also underlie the design of \texttt{dplyr} and the
Bioconductor infrastructure.

@@

The \texttt{plyranges} package can be obtained via the Bioconductor
project website \url{https://bioconductor.org} or accessed via Github
\url{https://github.com/sa-lee/plyranges}.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-availability}{%
\subsection{Data Availability}\label{data-availability}}

The BigWig file for the H3K27Me3 primary T CD8+ memory cells from
peripheral blood ChIP-seq data was downloaded from the
\texttt{AnnotationHub} package (2.12.0) under accession AH33458. The BAM
file corresponding to the H1 cell line ChIP-seq data is available at GEO
under accession GSM433167. The SNP array data for the H1 cell line data
is available at GEO under accession GSM1463263.

\hypertarget{software-versions}{%
\subsection{Software Versions}\label{software-versions}}

To produce the workflows as described in results section we used R
version 3.5 with the development version of \texttt{plyranges} (1.1.10)
and the \texttt{BioStrings} package (2.48.0) installed.

All code required to reproduce this article is available at
\url{https://github.com/sa-lee/plyranges-paper}.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages for
the code review they performed. This article was written with
\texttt{knitr} \cite{R-knitr} and the figures were made with
\texttt{ggbio} \cite{R-ggbio}.

>>>>

@@

\title{plyranges: a grammar of genomic data transformation  }

\author{\Large Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Dianne Cook \(^{\ref{monash}}\)\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author \href{mailto:michael.lawerence@gene.com}{michael.lawerence@gene.com} }\vspace{0.05in} \newline\normalsize\emph{}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia} () and Dianne Cook \(^{\ref{monash}}\) () and Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author \href{mailto:michael.lawerence@gene.com}{michael.lawerence@gene.com} } ()},
             pdfkeywords = {Bioconductor, Grammar, Genomes, Data Analysis},  
            pdftitle={plyranges: a grammar of genomic data transformation},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Dianne Cook \(^{\ref{monash}}\)} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author \href{mailto:michael.lawerence@gene.com}{michael.lawerence@gene.com} }} \hskip 15pt \emph{\small }

@@

\noindent There is a cognitive load placed on users in learning a data abstraction
from the Bioconductor project and understanding its appropriate use.
Users must navigate these abstractions to perform a genomic analysis
task, when a single data abstraction, a GRanges object will suffice. By
recognizing that the GRanges class follows `tidy' data principles, we
create a grammar of genomic data transformation, defining verbs for
performing actions on and between genomic interval data and providing a
way of performing common data analysis tasks through a coherent
interface to existing Bioconductor infrastructure. We implement this
grammar as a Bioconductor/R package called \texttt{plyranges}.

@@

Recently, there has been a concerted, community effort to standardize R
data structures and workflows around the notion of tidy data
\cite{Wickham2014-jc}. A tidy dataset is defined as a tabular data
structure that has observations as rows and columns as variables, and
all measurements pertain to a single observational unit. The tidy data
pattern is useful because it allows us to see how the data relate to the
design of an experiment and the variables measured. The \texttt{dplyr}
package \cite{Wickham2017-dplyr} defines an application programming
interface (API) that maps notions from the general relational algebra to
verbs that act on tidy data. These verbs can be composed together on one
or more tidy datasets with the pipe operator from the \texttt{magrittr}
package \cite{R-magrittr}. Taken together these features enable a user
to write human readable analysis workflows.

@@

{\centering \includegraphics{./diagrams/peak-viz-1}

@@

{\centering \includegraphics{./diagrams/titv-viz-1}

@@

The design of \texttt{plyranges} adheres to well understood principles
of language and API design: cognitive consistency, cohesion,
endomorphism and expressiveness \cite{Green1996-qg}. To varying degrees,
these principles also underlie the design of \texttt{dplyr} and the
Bioconductor infrastructure.

@@

\hypertarget{declarations}{%
\section{Declarations}\label{declarations}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages for
the code review they performed and users who have submitted feedback and
pull requests.

\hypertarget{availabliity-of-data-and-materials}{%
\subsection{Availabliity of data and
Materials}\label{availabliity-of-data-and-materials}}

The BigWig file for the H3K27Me3 primary T CD8+ memory cells from
peripheral blood ChIP-seq data from the Human Roadmap Epigenomics
project\\
was downloaded from the \texttt{AnnotationHub} package (2.13.1) under
accession AH33458 \cite{R-ahub, Roadmap-Epigenomics-Consortium2015-pr}.
The BAM file corresponding to the H1 cell line ChIP-seq data is
available at NCBI GEO under accession
\href{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM433167}{GSM433167}
\cite{ncbi-geo, Roadmap-Epigenomics-Consortium2015-pr}. The SNP array
data for the H1 cell line data is available at NCBI GEO under accession
\href{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GPL18952}{GPL18952}
\cite{Roadmap-Epigenomics-Consortium2015-pr}.

The \texttt{plyranges} package is open source under an Artistic 2.0
license \cite{R-plyranges-z}. The version (1.1.10) used in the paper is
archived under DOI:
\href{https://doi.org/10.5281/zenodo.1469841}{10.5281/zenodo.1469841}.
The software can be obtained via the Bioconductor project website
\url{https://bioconductor.org} or accessed via Github
\url{https://github.com/sa-lee/plyranges}.

To produce the workflows as described in results section we used R
version 3.5 with the development version of \texttt{plyranges}
(1.1.10)\\
and the \texttt{BioStrings} package (2.49.0) installed
\cite{r-core, R-plyranges-z, R-biostrings}.

This article was written with \texttt{knitr} \cite{R-knitr} and the
figures were made with \texttt{ggbio} \cite{R-ggbio}. All code required
to reproduce this article is available at
\url{https://github.com/sa-lee/plyranges-paper}.

\hypertarget{funding}{%
\subsection{Funding}\label{funding}}

SL is funded by an Australian Government Research Training Program (RTP)
scholarship and receives an additional top up scholarship from CSL.
Funding bodies had no role in data collection, analysis, and
interpretation.

\hypertarget{authors-contributions}{%
\subsection{Authors Contributions}\label{authors-contributions}}

SL, DC, ML contributed ideas towards the software. SL wrote the
manuscript and analysed data. SL and ML wrote the software. DC and ML
supervised the research and revised the manuscript. All authors have
read and approved the final version of the manuscript.

\hypertarget{ethics-approval-and-consent-to-participate}{%
\subsection{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}}

Not applicable to the current manuscript.

\hypertarget{competing-interests}{%
\subsection{Competing Interests}\label{competing-interests}}

Not applicable to the current manuscript.
